\textbf{Matching the forecast horizon with the relevant spatial and temporal processes and data sources}

### 2020-07-07 doc/response_2_reviews_Ecography3.txt

03-Jul-2020

Dear Dr. Adler:

Manuscript ID ECOG-05271.R1 entitled "Matching the forecast horizon with the relevant spatial and temporal processes and data sources" which you submitted to Ecography, has been reviewed.  The comments of the reviewer(s) and recommendation by the Subject Editor are included at the bottom of this letter.

The reviewer(s) and the Subject Editor have recommended some minor revisions to your manuscript.  Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript.

Recommendation by the Subject Editor (anonymous):
Your manuscript was evaluated by the reviewer who was the more critical reviewer of the original submission (the more positive reviewer was satisfied already with the revisions made earlier).

This was not an easy decision to make: reviewer 2 recommends rejection, because the same fundamental issues remain that were present basically in the original submission. Based on my own reading, I agree with the reviewer that your study has substantial limitations. At the same time, based on my own reading, I consider that the manuscript makes a very important point, and that it is generally very nicely written. Thus, in my view the models presented are simply "proofs of concepts" that help to discuss the (quite obvious but yet maybe not sufficiently acknowledged) fact that time-series models should do better in short-term predictions and space-for-time models should not better in long-term predictions. Given that making the examples more realistic and general would lead to a completely new manuscript, my options were basically to reject the manuscript, or accept it with these limitations. Given the above, I ended up to a decision that was favorable for your side. Thus, I invite you to prepare a minor revision, where you make one more effort to double-check that you don't overstate the generality of the findings, but rather present your models as proofs of concepts that help to illustrate your verbal argument. Related to this, I agree with the reviewer that you may have overstated the value of the simplistic linear weighting approach (even if you discuss its limitations nicely in the main text), as based on your text some readers might think that "just by figuring out the weighting functions one can make accurate predictions for any times-scale". This is of course not expected to be the case for any real case study, even if the optimal weighting would be known. So I would propose that you discuss even more critically the prospects of the weighted-model approach in the revised version.

RESPONSE: We thank the editor for their favorable decision, and for the opportunity to revise. We have addressed the reasonable concern about the potential value of the model weighting approach by adding a new, short paragraph to the Discussion: "On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems. The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}. The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data."

RESPONSE CONT'D: We also added caveats to the section at the end of the Introduction where we briefly describe our findings (changes tracked in red font).

Reviewer(s)' Comments to Author:
Reviewer: 2

Comments to the Author
This is a resubmission. It still suffers from the same problem as the previous version - the authors are not willing to explore the methods fully. The authors state they find the weighting interesting, but even this is only explored under one set of conditions. How do we know it will act the same if (for example) there is no change in climate?

RESPONSE: If there is no change in climate, there will be no need for model weighting. As our results show, the time series approach will work fine on its own.  Pointing this out seems unnecessary because the premise of the whole exercise was to predict the impact of environmental change.

I don't agree that the shift in weighting is so important, and I can't see that it is general. Only one specific scenario is used, so what will happen under different scenarios? And what if different models were used for the projections, e.g. adding a quadratic term?

RESPONSE: We are sorry the reviewer sees so little value in the analysis. However, it is not strictly accurate to say that we only explored one scenario. We studied two different model systems, and, at the reviewers' request, we also explored a non-stationary scenario. Perhaps the reviewer believes we are proposing a solution to a problem. Our real objective is to highlight a previously under-appreciated problem and propose one potential, and novel, solution. We believe we have added enough caveats and qualifiers to make this clear.

RESPONSE: As for the comment about adding a quadratic term, the spatial models for both the metacommunity example and the eco-evo example do include a quadratic effect of temperature (Eq. 1 in Appendix A, also shown in Figs. 2A and 4B). While the temporal model for the metacommunity example does not contain a quadratic effect of temperature (Appendix A Eq. 2, and Fig. 2B), the temporal model for the eco-evo example does (Appendix A Eq. 3). We added text in Appendix A to clarify this distinction. The reason the temporal model for the metacommunity example does not include a quadratic term is that the data from the baseline period show a linear relationship between temperature and biomass (Fig. 2B). Even if we fit a quadratic term, the coefficient would be close to zero and would not improve long-term forecasts. This illustrates a key point of the paper: different kinds of data represent different processes operating at different time scales.

A minor point: check that the descriptions of the models match the code. I was wondering why there isn't  a quadratic term on K(t) in the temporal model for biomass, but according to the code there isn't one on the spatial model either.

RESPONSE: We address this comment in the paragraph above. We double-checked our computer code and it is correct. The reviewer may have been looking at the spatial model for total biomass in the community turnover example, rather than the biomass of the focal species. That model does not include a quadratic term because the relationship is monotonic increasing. To minimize the chance of this confusion, we moved this block of code to the end of the script to follow the order in which results are presented.

### 2020-07-07 doc/manuscript_acceptedJuly2020.tex
final accepted version

\noindent \emph{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models. PBA wrote the first draft of the manuscript, and all authors contributed substantially to revisions.

\noindent \emph{Acknowledgements:} We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper. PBA was supported by the National Science Foundation (DEB-1927282) and the Utah Agriculture Experiment Station (grant 1322). EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative (GBMF4563).

@@

\noindent \textbf{Abstract:}   Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients. However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change. Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two. We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data. First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models. Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average. We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models. We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death. The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection. The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly. A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.

\section*{Abstract} 
Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients. However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change. Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two. We use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data. First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models. Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average. We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models. We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death. The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection. The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly. A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.

@@

Here we use two simulated case studies to 1) demonstrate why time-series and space-for-time approaches can make different predictions, 2) propose that the best model-building approaches for ecological forecasting 
may depend on the time horizon of the forecast, and 3) explore how time-series and space-for-time approaches might be 
combined via weighted averaging to make better forecasts at intermediate time scales. The first case study
focuses on how interspecific interactions affect the population dynamics of a focal species, and the second focuses on
an eco-evolutionary scenario. Our \comment{simulations illustrate} that: 
\begin{enumerate}
	\item For short-term forecasts, phenomenological time-series approaches \comment{may be} hard to beat, whereas longer-term forecasts \comment{may} require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and fast responses of current assemblages to interannual environmental variation, while data spanning spatial gradients capture the long-term outcome of interactions between fast and slow processes.  Whether predictive models should be trained using longitudinal or spatial data sets, or both, \comment{likely} depends on the time-scale of the desired forecast.
	\item A key challenge for future research is determining the rate at which slow processes begin to influence dynamics.
\end{enumerate}

@@

\comment{On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems. The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}. The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data.}

@@

@@

The temporal model, which we call $T$, starts with a time-series of ``observed" population sizes, or total biomasses, at one location, $N(t)$, for $t=1...n$ (the spatial index is suppressed because we only focus on one location at a time). In the community turnover example, we fit the following regression, which predicts biomass at time $t+1$ as a function of biomass ($N(t)$) and annual temperature ($K(t)$) at time $t$,
\begin{equation}
\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
\label{eqn:temporal_regression_community1}
\end{equation}
In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, \comment{which includes a quadratic effect of temperature,} is 
\begin{equation}
\ln\left(\frac{N(t+1)}{N(t)}\right) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
\label{eqn:temporal_regression_ecoevo2}
\end{equation}
This version of the temporal model returns a per capita growth rate on the log scale. To predict population size at the next time step, we exponentiate the growth rate and multiply it by the current population size: $\exp(T(N(t),K(t))) N(t)$.

>>>>

@@

\noindent \textbf{Abstract:}   Most phenomenological, statistical models used to generate ecological forecasts take either a time-series approach, based on long-term data from one location, or a space-for-time approach, based on data describing spatial patterns across environmental gradients. However, the magnitude and even the sign of environment-response relationships detected using these two approaches often differs, leading to contrasting predictions about responses to future environmental change. Here we consider how the forecast horizon determines whether more accurate predictions come from the time-series approach, the space-for-time approach, or a combination of the two. As proof of concept, we use simulated case studies to show that forecasts for short and long forecast horizons need to focus on different ecological processes, which are reflected in different kinds of data. First, we simulated population or community dynamics under stationary temperature using two simple, mechanistic models. Second, we fit statistical models to the simulated data using a time-series approach, a space-for-time approach, or a weighted average. We then forecast the response to a temperature increase using the statistical models, and compared these forecasts to temperature effects simulated by the mechanistic models. We found that the time-series approach made accurate short-term predictions because it captured initial conditions and effects of fast processes such as birth and death. The space-for-time approach made more accurate long-term predictions because it better captured the influence of slower processes such as evolutionary and ecological selection. The weighted average made accurate predictions at all time scales, including intermediate time-scales where the other two approaches performed poorly. A weighted average of time-series and space-for-time approaches shows promise, but making this weighted model operational will require new research to predict the rate at which slow processes begin to influence dynamics.

@@

Here we use two simulated case studies to 1) demonstrate why time-series and space-for-time approaches can make different predictions, 2) propose that the best model-building approaches for ecological forecasting 
may depend on the time horizon of the forecast, and 3) explore how time-series and space-for-time approaches might be 
combined via weighted averaging to make better forecasts at intermediate time scales. The first case study
focuses on how interspecific interactions affect the population dynamics of a focal species, and the second focuses on
an eco-evolutionary scenario. Our simulations illustrate that: 
\begin{enumerate}
	\item For short-term forecasts, phenomenological time-series approaches may be hard to beat, whereas longer-term forecasts may require accounting for the influence of slow processes such as evolutionary and ecological selection as well as dispersal.
	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and fast responses of current assemblages to interannual environmental variation, while data spanning spatial gradients capture the long-term outcome of interactions between fast and slow processes.  Whether predictive models should be trained using longitudinal or spatial data sets, or both, likely depends on the time-scale of the desired forecast.
	\item A key challenge for future research is determining the rate at which slow processes begin to influence dynamics.
\end{enumerate}

@@

On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems. The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}. The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data.

@@

\subsubsection*{Statement of authorship:} PBA and EPW designed the study, PBA and MHC built the models, and PBA analyzed the models. PBA wrote the first draft of the manuscript, and all authors contributed substantially to revisions.

\subsubsection*{Acknowledgements:} We thank Heather Lynch and Juan Manuel Morales for comments that improved early drafts of the paper. PBA was supported by the National Science Foundation (DEB-1927282) and the Utah Agriculture Experiment Station (grant 1322). EPW was supported by the Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative (GBMF4563).

@@

The temporal model, which we call $T$, starts with a time-series of ``observed" population sizes, or total biomasses, at one location, $N(t)$, for $t=1...n$ (the spatial index is suppressed because we only focus on one location at a time). In the community turnover example, we fit the following regression, which predicts biomass at time $t+1$ as a function of biomass ($N(t)$) and annual temperature ($K(t)$) at time $t$,
\begin{equation}
\ln(N(t+1)) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +  \varepsilon
\label{eqn:temporal_regression_community1}
\end{equation}
In the eco-evolutionary example, the response variable is the log of the population growth rate. The regression, which includes a quadratic effect of temperature, is 
\begin{equation}
\ln\left(\frac{N(t+1)}{N(t)}\right) = T(N(t),K(t)) = \beta^T_0 +  \beta^T_1 \ln(N(t)) +\beta^T_2 K(t)  +\beta^T_3 K(t)^2 +  \varepsilon
\label{eqn:temporal_regression_ecoevo2}
\end{equation}
This version of the temporal model returns a per capita growth rate on the log scale. To predict population size at the next time step, we exponentiate the growth rate and multiply it by the current population size: $\exp(T(N(t),K(t))) N(t)$.
